---
title: "Meta-analysis"
output:
  pdf_document:
    df_print: paged
date: '2022-07-04'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=F)
```

```{r}
library(pacman)
p_load(
  meta,
  here, 
  esc,
  tidyverse,
  pwr
)

```

```{r}
# function that pools two different groups
pool.groups = function(n1, n2, m1, m2, sd1, sd2) {

    n1 = n1
    n2 = n2
    m1 = m1
    m2 = m2
    sd1 = sd1
    sd2 = sd2

    if (is.numeric(n1) == FALSE) {
        stop("'n1' must by of type numeric().")
    }

    if (is.numeric(n2) == FALSE) {
        stop("'n2' must by of type numeric().")
    }

    if (n1 < 1 | n2 < 1) {
        stop("'n1' and 'n2' must both the greater than 0.")
    }

    if (is.numeric(m1) == FALSE) {
        stop("'m1' must by of type numeric().")
    }

    if (is.numeric(m2) == FALSE) {
        stop("'m2' must by of type numeric().")
    }

    if (is.numeric(sd1) == FALSE) {
        stop("'sd1' must by of type numeric().")
    }

    if (is.numeric(sd2) == FALSE) {
        stop("'sd2' must by of type numeric().")
    }

    Npooled = n1 + n2
    Mpooled = (n1 * m1 + n2 * m2)/(n1 + n2)
    SDpooled = sqrt(((n1 - 1) * sd1^2 + (n2 - 1) * sd2^2 + (((n1 * n2)/(n1 + n2)) * (m1^2 + m2^2 - 2 * m1 *
        m2)))/(n1 + n2 - 1))

    return(data.frame(Mpooled, SDpooled, Npooled))
}
```

## Step 1:

Combine different languages within the same study and language type (tonal or non-tonal) using the pooling formula below on the raw metrics

```{r}
meta_original <- read.csv(here("data/meta-analysis", "meta-analysis_original.csv"))
#separating language column:
meta_original[c('Tonal_language', 'Comparison_language')] <- str_split_fixed(meta_original$Language.comparison, ' vs. ', 2)
meta_original <- subset(meta_original, select = -Language.comparison)

meta_original$Measure <- trimws(meta_original$Measure)
meta_original$Study <- trimws(meta_original$Study)
meta_original$Tonal_language <- trimws(meta_original$Tonal_language)
meta_original$Comparison_language <- trimws(meta_original$Comparison_language)
# trim out white spaces
```

```{r}
# function that pools language groups
combine_lang <- function(data, study, measure, measure_overall, lang1, lang2, Tonal = T) {
  dat <- data %>%
    filter(Study == study) %>%
    filter(Measure == measure) %>%
    select(Tonal_N, Tonal.Mean, Tonal.SD)

  if(Tonal == F) {
    dat <- data %>%
      filter(Study == study) %>%
      filter(Measure == measure) %>%
      select(Non_tonal_N, Non.tonal.Mean, Non.tonal.SD)
  }

  name <- c("n1", "n2", "m1", "m2", "sd1", "sd2")
  values <- c(dat$Tonal_N, dat$Tonal.Mean, dat$Tonal.SD)

   if(Tonal == F) {
    values <- c(dat$Non_tonal_N, dat$Non.tonal.Mean, dat$Non.tonal.SD)
  }

  for(num in c(1:6)){
  assign(name[num], values[num])
  }

  result<- pool.groups(n1, n2, m1, m2, sd1, sd2)

  if(Tonal){
    data <- data[-which(data$Study == study & data$Measure ==measure & data$Tonal_language == lang2),]
    data$Tonal_N[which(data$Study == study & data$Measure ==measure)] <-   result$Npooled
    data$Tonal_N[which(data$Study == study & data$Measure == measure_overall)] <- result$Npooled # updating the general task n
    data$Tonal.Mean[which(data$Study == study & data$Measure ==measure)] <- result$Mpooled
    data$Tonal.SD[which(data$Study == study & data$Measure ==measure)] <- result$SDpooled
    data$Tonal_language[which(data$Study == study & data$Measure ==measure)] <- paste0(lang1, " or ", lang2)
    data$Tonal_language[which(data$Study == study & data$Measure ==measure_overall)] <- paste0(lang1, " or ", lang2) # updating general task language
  }

  if(Tonal == F){
    data <- data[-which(data$Study == study & data$Measure ==measure & data$Comparison_language == lang2),]
    data$Non_tonal_N[which(data$Study == study & data$Measure ==measure)] <-   result$Npooled
    data$Non_tonal_N[which(data$Study == study & data$Measure == measure_overall)] <-   result$Npooled
    data$Non.tonal.Mean[which(data$Study == study & data$Measure ==measure)] <- result$Mpooled
    data$Non.tonal.SD[which(data$Study == study & data$Measure ==measure)] <- result$SDpooled
    data$Comparison_language[which(data$Study == study & data$Measure ==measure)] <- paste0(lang1, " or ", lang2)
    data$Comparison_language[which(data$Study == study & data$Measure ==measure_overall)] <- paste0(lang1, " or ", lang2)

  }

  return(data)
}
```

```{r}
# Bradley (2016), Mandarin and Yoruba, contour, interval, and register:
meta_original <- meta_original %>% #removing the general task row for Yoruba
  filter(!(Study == "Bradley (2016)" & Measure == "Melodic comparison" & Tonal_language == "Yoruba"))
meta_original <- combine_lang(meta_original, study = "Bradley (2016)", measure = "Contour", lang1 = "Mandarin", lang2 = "Yoruba", measure_overall = "Melodic comparison")
meta_original <- combine_lang(meta_original, study = "Bradley (2016)", measure = "Interval", lang1 = "Mandarin", lang2 = "Yoruba", measure_overall = "Melodic comparison")
meta_original <- combine_lang(meta_original, study = "Bradley (2016)", measure = "Register", lang1 = "Mandarin", lang2 = "Yoruba", measure_overall = "Melodic comparison")

#Zheng & Samuel 2018, English and Korean,
meta_original<- combine_lang(meta_original, study = "Zheng \\& Samuel (2018)", measure = "Melody comparison", lang1 = "Korean", lang2 = "English", Tonal = F, measure_overall = NULL)
meta_original<- combine_lang(meta_original, study = "Zheng \\& Samuel (2018)", measure = "Pitch precision", lang1 = "Korean", lang2 = "English", Tonal = F, measure_overall = NULL)

# Hove et al. 2010 Chinese and Hmong 
meta_original<- combine_lang(meta_original, study = "Hove et al. (2010)", measure = "Relative pitch recognition", lang1 = "Chinese", lang2 = "Hmong", Tonal = T, measure_overall = NULL)

# Jasmin et al. 2021 English and Spanish 
meta_original<- combine_lang(meta_original, study = "Jasmin et al. (2021)", measure = "Pitch Discrimination", lang1 = "English", lang2 = "Spanish", Tonal = F, measure_overall = NULL)
meta_original<- combine_lang(meta_original, study = "Jasmin et al. (2021)", measure = "Duration Discrimination", lang1 = "English", lang2 = "Spanish", Tonal = F, measure_overall = NULL)
```

```{r}
meta_grouped <- meta_original
```

## Step 2:

calculate cohen's d, hedge's g and their SE

```{r}
sd_pooled <- function(n1, n2, s1, s2) { # function that calculates the pooled sd of tonal and non-tonal for each study/entry
  sd = sqrt(
    ((n1-1)*s1^2 + (n2-1)*s2^2)/
      ((n1-1) + (n2-1))
    )
  return(sd)
}

se_smd <- function(n1, n2, smd) {
  se <- sqrt(
    (n1+n2)/(n1*n2) + smd^2/(2*(n1 + n2))
  )
}
```

```{r}
meta_grouped$sd_pool <- sd_pooled(n1 = meta_grouped$Tonal_N, n2 = meta_grouped$Non_tonal_N, s1 = meta_grouped$Tonal.SD, s2 = meta_grouped$Non.tonal.SD)

meta_grouped$cohen_d <- (meta_grouped$Tonal.Mean-meta_grouped$Non.tonal.Mean)/meta_grouped$sd_pool

meta_grouped$cohen_d <- ifelse(is.na(meta_grouped$cohen_d), meta_grouped$Cohen.s.D, meta_grouped$cohen_d) # add Alexander2008 to cohen's d list; this study reported cohen's d directly 

meta_grouped$total_n <- meta_grouped$Tonal_N + meta_grouped$Non_tonal_N

meta_grouped$hedge_g <- hedges_g(meta_grouped$cohen_d, meta_grouped$total_n)
```

```{r}
meta_grouped$se_smd <- se_smd(n1 = meta_grouped$Tonal_N, n2 = meta_grouped$Non_tonal_N, smd = meta_grouped$cohen_d)
meta_grouped$se_g <- se_smd(n1 = meta_grouped$Tonal_N, n2 = meta_grouped$Non_tonal_N, smd = meta_grouped$hedge_g)

esc_mean_sd(grp1m = 0.8060976, grp2m = 0.6600000,
            grp1sd = 0.18554480, grp2sd = 0.2000000,
            grp1n = 41, grp2n = 26) # checking that formula is coded correctly using Melodic comparison from Bradley (2016) as an example
```

## Step 3: Combine subtasks into composite effect sizes

Next, we combine subtasks effect sizes into a total effect size for studies that broke down results at subtask levels (e.g., 5 vs. 10 cents difference)

Calculation method comes from Page 228 of Borenstein, M., Hedges, L. V., Higgins, J. P. T., & Rothstein, H. R. (2009). Introduction to meta-analysis. John Wiley & Sons

The correlation is assumed to be 1 between these subtasks

```{r}
sum_calc <- function(vec){
  sum = 0
  for (i in c(1:length(vec))){
  for(j in c(1:length(vec))){
    if(j == i){
      sum = sum
    }
    else{
      sum = sum + sqrt(vec[i]*vec[j])
    }
    j = j+1
  }
  i = i + 1
  }
  return(sum)
}

var_combine <- function(vector){
  m = length(vector)
  sum = sum(vector)
  sum_combine = sum_calc(vec = vector)
  var = (1/m)**2*(sum + sum_combine)
  se = sqrt(var)
  return(c(variance = var, se = se))
}
```

```{r}
meta_processed <- meta_grouped
meta_processed$var_smd <- meta_processed$se_smd**2
meta_processed$var_g <- meta_processed$se_g**2
```

```{r}
combine_subscore <- function(data, study, measure, specific_measures = NULL){
  dat <- data %>%
    filter(Study == study & Measure.overall == measure) %>%
    filter(Measure != measure)

  if(!is.null(specific_measures)){
    dat <- data %>%
      filter(Study == study & Measure %in% specific_measures)
  }

  g.mean = mean(dat$hedge_g)
  g.var_se <- var_combine(dat$var_g)

  smd.mean = mean(dat$cohen_d)
  smd.var_se <- var_combine(dat$var_smd)

  result <- c(g.mean = g.mean, g = g.var_se, smd.mean = smd.mean, smd = smd.var_se)

  data$hedge_g[which(data$Study == study & data$Measure == measure)] <- result["g.mean"]
  data$se_g[which(data$Study == study & data$Measure == measure)] <- result["g.se"]
  data$var_g[which(data$Study == study & data$Measure == measure)] <- result["g.variance"]

  data$cohen_d[which(data$Study == study & data$Measure == measure)] <- result["smd.mean"]
  data$se_smd[which(data$Study == study & data$Measure == measure)] <- result["smd.se"]
  data$var_smd[which(data$Study == study & data$Measure == measure)] <- result["smd.variance"]

  if(is.null(specific_measures)){
    data <- data %>%
    filter(!(Study == study & Measure.overall == measure & Measure != measure ))
  }

  else{
     data <- data %>%
      filter(!(Study == study & Measure %in% specific_measures))
  }


  return(data)
}
```

```{r}
# Bent et al. (2006) Nonspeech pitch discrimination:
meta_processed <- combine_subscore(meta_processed, "Bent et al. (2006)", "Nonspeech pitch discrimination")

# Bradley 2016 Melodic comparison:
meta_processed <- combine_subscore(meta_processed, "Bradley (2016)", "Melodic comparison")

# Bidelman et al. (2011) Chordal detuning:
meta_processed <- combine_subscore(meta_processed, "Bidelman et al. (2011)", "Chordal detuning discrimination")

# Bidelman et al. (2013) Melody discrimination:
meta_processed <- combine_subscore(meta_processed, "Bidelman et al. (2013)", "Melody discrimination")

# Giuliano et al. (2011) Pitch discrimination & Interval discrimination:
meta_processed <- combine_subscore(meta_processed, "Giuliano et al. (2011)", "Pitch discrimination")
meta_processed <- combine_subscore(meta_processed, "Giuliano et al. (2011)", "Interval discrimination")

# Ngo et al. (2016) Montreal Battery of Evaluation of Amusia
meta_processed <- combine_subscore(meta_processed, "Ngo et al. (2016)", "Montreal Battery of Evaluation of Amusia", specific_measures = c("Scale", "Contour", "Interval"))

# Peretz et al. (2011) Montreal Battery of Evaluation of Amusia & Pitch change detection
meta_processed <- combine_subscore(meta_processed, "Peretz et al. (2011)", "Montreal Battery of Evaluation of Amusia", specific_measures = c("Scale", "Contour", "Interval"))
meta_processed <- combine_subscore(meta_processed, "Peretz et al. (2011)", "Pitch change detection")

# Chen et al. (2016) Montreal Battery of Evaluation of Amusia
meta_processed <- combine_subscore(meta_processed, "Chen et al. (2016)", "Montreal Battery of Evaluation of Amusia", specific_measures = c("Scale", "Contour", "Interval"))
```

```{r}
meta_processed$Measure <-replace(meta_processed$Measure, meta_processed$Measure == "Montreal Battery of Evaluation of Amusia", "Melody combined") # replacing MBEA with melody combined since we only combined the melody subtasks
```

```{r}
# Chen 2016: Reported both accuracy and raw score. Accuracy and raw score not 100% consistent (maybe due to excluding trials). Will calculate using accuracy since it takes into account trial-excluding
meta_processed <- meta_processed %>%
  filter(!(Study == "Chen et al. (2016)" & Measure == "Melody subtest"))
```

```{r}
classification <- read.csv(here("data/meta-analysis", "Measure_classification.csv"))

meta_processed <- full_join(meta_processed, classification) %>%
  distinct(.)
```

## Step 4

```{r}
# Removing rhythm tests, meter, memory, speed, and non-traditional (nonspeech pitch contour identification) tests for now
meta_filtered <- meta_processed %>%
  filter(!Measure %in% c("Pitch processing speed", "Pitch memory", "Nonspeech pitch contour identification", "Meter", "Memory"))

meta_filtered <- meta_filtered %>%
  mutate(hedge_g = case_when(
    Flip == T ~ -1*hedge_g, # tasks with flip == T are those that are reverse coded (high value == low performance)
    TRUE~hedge_g))
```

## Step 5 pooling effects-all study measures

```{r}
meta_processed_final <- meta_filtered %>%
  filter(Study != "Zhang et al. (2020)") # removing study comparing Japanese and Mandarin
```

```{r}
meta_processed_final$Study_Measure <- paste0(meta_processed_final$Study, ": ",  meta_processed_final$Measure)

# removing the extra backslash 
meta_processed_final <- meta_processed_final %>%
  mutate(
    # this complexity is needed because it is treated as an escaped character
    Study_Measure = str_replace_all(Study_Measure, "[^[A-Za-z,0-9]()\\-\\.:& //]", "")
  )
```

```{r}
meta_processed_final %>%
  count(Measure.Classification) # number of tasks per category
```

## Melodic tasks 

```{r fig.width = 10}
meta_melodic <- meta_processed_final %>%
                   filter(Measure.Classification == "Melodic") 

melodic.gen.random  <- metagen(TE = hedge_g,
                 seTE = se_g, 
                 studlab = Study_Measure,
                 data = meta_melodic,
                 sm = "SMD",
                 fixed = FALSE,
                 random = TRUE,
                 method.tau = "REML",
                 hakn = TRUE,
                 title = "tonelang")

melodic.gen.random$Power <- pwr.t2n.test(n1= meta_melodic$Tonal_N, n2 = meta_melodic$Non_tonal_N, sig.level = 0.05, d = 0.50, power = NULL)$power

forest.melodic <- forest.meta(melodic.gen.random,
            sortvar = TE)
```

## Fine-tuned tasks

```{r fig.width = 10}
meta_pitch <- meta_processed_final %>%
                filter(Measure.Classification == "Fine-tuned")

pitch.gen.random  <- metagen(TE = hedge_g,
                 seTE = se_g, #
                 studlab = Study_Measure,
                 data = meta_pitch,
                 sm = "SMD",
                 fixed = FALSE,
                 random = TRUE,
                 method.tau = "REML",
                 hakn = TRUE,
                 title = "tonelang")

pitch.gen.random$Power <- pwr.t2n.test(n1= meta_pitch$Tonal_N, n2 = meta_pitch$Non_tonal_N, sig.level = 0.05, d = 0.5, power = NULL)$power

forest.pitch <- forest.meta(pitch.gen.random,
            sortvar = TE)
```

## Rhythm:

```{r fig.width = 10}
meta_rhythm <- meta_processed_final %>%
                filter(Measure.Classification == "Rhythm")

rhythm.gen.random  <- metagen(TE = hedge_g,
                 seTE = se_g, #
                 studlab = Study_Measure,
                 data = meta_rhythm,
                 sm = "SMD",
                 fixed = FALSE,
                 random = TRUE,
                 method.tau = "REML",
                 hakn = TRUE,
                 title = "tonelang")

rhythm.gen.random$Power <- pwr.t2n.test(n1= meta_rhythm$Tonal_N, n2 = meta_rhythm$Non_tonal_N, sig.level = 0.05, d = 0.50, power = NULL)$power

forest.rhythm <- forest.meta(rhythm.gen.random,
            sortvar = TE)
```

## Numbers from prior studies 

```{r}
study_num <- meta_processed %>%
  select(Study, Tonal_N, Non_tonal_N) %>%
  distinct(.) %>%
  pull(Tonal_N, Non_tonal_N)
```

```{r}
study_total_num <- meta_processed_final %>%
  select(Study) %>%
  distinct(.) %>%
  nrow()
```

```{r}
# save(list = c("melodic.gen.random", "pitch.gen.random", "rhythm.gen.random", "study_num", "study_total_num"), file = here("results", "meta_analyses.RData"))
# uncomment to save 
```

## Median power of analyzed studies 
```{r}
round(median(melodic.gen.random$Power),2)
round(median(pitch.gen.random$Power),2)
round(median(rhythm.gen.random$Power),2)
```

## Number of tonal speakers analyzed 
```{r}
meta_original <- read.csv(here("data/meta-analysis", "meta-analysis_original.csv"))

meta_original %>%
  filter(Study %in% meta_processed_final$Study) %>%
  select(Study, Tonal_N) %>%
  distinct(.) %>%
  pull(Tonal_N) %>%
  sum() # total number of tonal participants analyzed 
```
